Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-12-12 Build_167/MDPI_submission_B167.tex
Commit 2023-12-13 01:11

We calculated the linear SDR trends, from the departures of the mean
daily values from the daily climatology and for the three sky conditions
(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
only the time series under all-sky conditions; the plots for clear-sky
and cloud-sky conditions, are very similar and are shown in the Appendix
(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
smaller trend for clear-skies (\(0.097\,\%/y\)) and a negative weaker
trend for cloudy-skies (\(0.41\,\%/y\)). In the studied period, there is
no significant break or change in the variability of the time series.
Other studies for the European region reported a change of the SDR
slope, around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years
before the start of our records. It is interesting to note, that for the
observations period, the trend of the TSI is \(-0.00022\,\%/y\), and
thus we can not attribute any major effect on SDR trend to Solar
variability.

@@

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky conditions & Trend [\%/year] & Trend p-value & Pearson cor. & Days\\
\midrule
All skies & 0.377 & 0 & 0.0913 & 10251\\
Clear skies & 0.097 & 0 & 0.1381 & 2689\\
Cloudy skies & 0.408 & 0 & 0.0810 & 4937\\
\bottomrule
\end{tabu}
\end{table}

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
\toprule
Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & slope.stat\_sig\\
\midrule
\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.31200} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
\cmidrule{1-7}
0.1460 & 0.178000 & 0.06170 & 0.2483832 & All skies & Spring & 82.2\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.09580} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
\cmidrule{1-7}
0.4300 & 0.006670 & 0.23500 & 0.4844650 & All skies & Autumn & 99.3\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.3570} & \cellcolor{gray!6}{0.007260} & \cellcolor{gray!6}{0.23000} & \cellcolor{gray!6}{0.4800278} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.3}\\
\cmidrule{1-7}
0.0317 & 0.721000 & 0.00445 & 0.0667338 & Clear skies & Spring & 27.9\\
\cmidrule{1-7}
\cellcolor{gray!6}{-0.0931} & \cellcolor{gray!6}{0.007770} & \cellcolor{gray!6}{0.22700} & \cellcolor{gray!6}{-0.4764646} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{99.2}\\
\cmidrule{1-7}
0.1100 & 0.308000 & 0.03710 & 0.1926085 & Clear skies & Autumn & 69.2\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.8250} & \cellcolor{gray!6}{0.000123} & \cellcolor{gray!6}{0.41500} & \cellcolor{gray!6}{0.6439160} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-7}
0.0942 & 0.425000 & 0.02210 & 0.1485860 & Cloudy skies & Spring & 57.5\\
\cmidrule{1-7}
\cellcolor{gray!6}{-0.0983} & \cellcolor{gray!6}{0.414000} & \cellcolor{gray!6}{0.02400} & \cellcolor{gray!6}{-0.1548507} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{58.6}\\
\cmidrule{1-7}
0.1950 & 0.460000 & 0.01960 & 0.1401398 & Cloudy skies & Autumn & 54.0\\
\bottomrule
\end{tabu}
\end{table}

>>>>

We calculated the linear SDR trends, from the departures of the mean
daily values from the daily climatology and for the three sky conditions
(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
only the time series under all-sky conditions; the plots for clear-sky
and cloud-sky conditions, are very similar and are shown in the Appendix
(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
smaller trend for clear-skies (\(0.097\,\%/y\)) and a negative weaker
trend for cloudy-skies (\(0.41\,\%/y\)). In the studied period, there is
no significant break or change in the variability of the time series.
Other studies for the European region reported a change of the SDR
slope, around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years
before the start of our records. It is interesting to note, that for the
observations period, the trend of the TSI is \(-0.00024\,\%/y\), and
thus we can not attribute any major effect on SDR trend to Solar
variability.

@@

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky conditions & Trend [\%/year] & Trend p-value & Pearson cor. & Days\\
\midrule
All skies & 0.376 & 0 & 0.0911 & 10250\\
Clear skies & 0.097 & 0 & 0.1381 & 2689\\
Cloudy skies & 0.408 & 0 & 0.0810 & 4937\\
\bottomrule
\end{tabu}
\end{table}

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
\toprule
Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & N\_eff & t\_eff & t\_eff\_cri & conf\_2.5 & conf\_97.5\\
\midrule
\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.31200} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{24.08252} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.063524} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
\cmidrule{1-11}
0.1360 & 0.219000 & 0.05150 & 0.2270307 & All skies & Spring & 44.44554 & 1e-07 & 2.014797 & -0.0011 & 0.0015253\\
\cmidrule{1-11}
\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.09580} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{23.37502} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.066823} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
\cmidrule{1-11}
0.4300 & 0.006670 & 0.23500 & 0.4844650 & All skies & Autumn & 21.43236 & 1e-07 & 2.077062 & -0.0011 & 0.0015253\\
\cmidrule{1-11}
\cellcolor{gray!6}{0.3570} & \cellcolor{gray!6}{0.007260} & \cellcolor{gray!6}{0.23000} & \cellcolor{gray!6}{0.4800278} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{18.55573} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.096421} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
\cmidrule{1-11}
0.0317 & 0.721000 & 0.00445 & 0.0667338 & Clear skies & Spring & 21.33345 & 0e+00 & 2.077636 & -0.0011 & 0.0015253\\
\cmidrule{1-11}
\cellcolor{gray!6}{-0.0931} & \cellcolor{gray!6}{0.007770} & \cellcolor{gray!6}{0.22700} & \cellcolor{gray!6}{-0.4764646} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{25.12298} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.059027} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
\cmidrule{1-11}
0.1100 & 0.308000 & 0.03710 & 0.1926085 & Clear skies & Autumn & 23.96832 & 0e+00 & 2.064043 & -0.0011 & 0.0015253\\
\cmidrule{1-11}
\cellcolor{gray!6}{0.8250} & \cellcolor{gray!6}{0.000123} & \cellcolor{gray!6}{0.41500} & \cellcolor{gray!6}{0.6439160} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{14.68318} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.135462} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
\cmidrule{1-11}
0.0942 & 0.425000 & 0.02210 & 0.1485860 & Cloudy skies & Spring & 35.40650 & 1e-07 & 2.029275 & -0.0011 & 0.0015253\\
\cmidrule{1-11}
\cellcolor{gray!6}{-0.0983} & \cellcolor{gray!6}{0.414000} & \cellcolor{gray!6}{0.02400} & \cellcolor{gray!6}{-0.1548507} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{37.66778} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.024981} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
\cmidrule{1-11}
0.1950 & 0.460000 & 0.01960 & 0.1401398 & Cloudy skies & Autumn & 28.41812 & 2e-07 & 2.047050 & -0.0011 & 0.0015253\\
\bottomrule
\end{tabu}
\end{table}

### 2023-12-12 appendix.tex
Commit 2023-11-08 21:10
Commit 2023-11-09 04:14
Commit 2023-11-09 05:11
Commit 2023-11-09 18:14
Commit 2023-12-13 01:11

%% optional
\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixsections{one} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A"). If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
\appendix
% \section{}

\section[\appendixname~\thesection]{} % test

\  \

% copy output from Article.tex

\begin{figure}[h!]

{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-5}}

\caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023. The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
\end{figure}

\  \

\begin{figure}[h!]

{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-8}}

\caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
\end{figure}

\begin{figure}[h!]

{\centering \includegraphics[width=1\linewidth]{./images/SzaTrendsSeasTogether-2}}

\caption{Long term trends of SDR as a function of SZA separately form morning and afternoon
periods, by season (rows of plots) for the three sky conditions (columns of plots).
Solid shapes represent statistical significant trends ($p<0.005$).
Cases where $p<0.005$ or with less than $85$ observations may be missing from view.}\label{fig:SZAtrendSeason}
\end{figure}

\ \

>>>>

% % \renewcommand\thefigure{\thesection.\arabic{figure}} %% my
% % \renewcommand\thefigure{\thesection.\arabic{figure}}
%
% %% From R rticle MDPI template
% %% optional
% \appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
% \appendixsections{one} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A"). If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
%
% % \appendix
% % \section{}
%
% \appendix
% \setcounter{secnumdepth}{0}
% \section{Appendix}
%
% \setcounter{figure}{0}    %% my

%% From raw MDPI template
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section[\appendixname~\thesection]{}

% copy output from Rmd

\begin{figure}[h!]
    {\centering
        \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-7}

    }
    \caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023. The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
\end{figure}

\begin{figure}[h!]
    {\centering
        \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-10}

    }
    \caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
\end{figure}

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering
            \includegraphics[width=1.0\linewidth]{./images/SzaTrendsSeasTogether-2}
        }
        \caption{Long term trends of SDR as a function of SZA separately form morning
               and afternoon periods, by season (rows of plots) for the three sky
               conditions (columns of plots).
               Solid shapes represent statistical significant trends ($p<0.005$).
               Cases where $p<0.005$ or with less than $85$ observations may be
               missing from view.}\label{fig:SZAtrendSeason}
    \end{adjustwidth}
\end{figure}

\FloatBarrier
